Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate

### 2015-07-21 masters_paper/paper.tex
Fixed bug with X_lrg for original knockoffs

\subsection{Knockoff Filter}
    Let us consider the consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
    \[\mathbf{y} = X\beta + \mathbf{z}\]
    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise. For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$. I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$. The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$. In the situation where $\beta$ is sparse and there is reason to expect that $x_j$ and $y$ are uncorrelated for many $j$, the knockoff filter is a method to select a subset of the $x_j$ which are likely correlated with $y$. \par
    The method specifically seeks to control the false discovery rate (FDR), which is the expected portion of some selection of $x_j$ which are, in fact, uncorrelated with $y$. For a procedure which chooses a subset of variables $\hat S\subseteq \{1,\ldots,p\}$, this is defined as  
    \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: \beta_j=0 \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
    FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$. \par

    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being less correlated with $y$. In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$, the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized. In other words, where $\hat \Sigma := X^TX$, 
    \[ \tilde X^T\tilde X=\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}\]
    for some vector $s\in \R^p$ such that, where $\diag{\Sigma}$ is the vector made from the main diagonal of $\Sigma$, $\diag{\Sigma} - \mathbf s$ is small. Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$. In the case where $X_j$ is uncorrelated with $\mathbf y$, $\tilde X_j$ will also be uncorrelated with $\mathbf y$. How these knockoffs are actually generated will be described shortly.\par

@@

    In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second. Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor. On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $\Sigma_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature. This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $\Sigma_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$. \par

@@

\subsection{Original Knockoff Features}
    The original formulation of the method offers two similar methods of constructing knockoffs. Both of these will, by construction, have exactly the property that
    \[ \tilde X^T\tilde X=\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}.\]
    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $X_L$ is
    \[ \Sigma_L= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} \Sigma & \Sigma - \diag{\mathbf s} \\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \]
    $A$, the Schur complement of $\Sigma$ in $\Sigma_L$ can be calculated as 
    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}\Sigma^{-1}\diag{\mathbf s} \]
    For $\Sigma_L$ to exist, $\Sigma_L$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper} 
    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2\Sigma - \diag{\mathbf s} \succeq 0 \]
    Given this is true, $A$ can be factored as $A=C^TC$. Combining this with a satisfactory $\mathbf s$ and an orthonormal matrix $\tilde U\in\R^{n\times p}$ such that $\tilde U^T X = 0$, a $\tilde X$ fulfilling the desired properties can be calculated as 
    \[ \tilde X = X(I-\Sigma^{-1}\,\diag{\mathbf s}) + \tilde UC\]
    As mentioned above, $\mathbf s$ should be chosen so as to make $\diag{\Sigma}-\mathbf s$ small. Since each $X_j$ has been normalized, this means that $\diag{\Sigma}=\mathbf 1$, so $\mathbf 1 - \mathbf s$ should be minimized in accordance with the restrictions on $\mathbf s$. The two methods differ in how $\mathbf s$ is chosen:
    \begin{itemize}
        \item \textit{Equi-correlated knockoffs}: Each original features is set to have the same correlation with its knockoff by setting $\mathbf s = 2\min\{\lambda_{min}(\Sigma),1\}\mathbf 1$. For all simulations in this paper, this was the method used.
        \item \textit{SDP knockoffs}: The $\mathbf s$ which minimizes the average correlation between knockoff and original features can be found via a semi-definite programing problem:
            \begin{center}
                \begin{tabular}{r l}
                    minimize & $\|\mathbf1-\mathbf s\|_1$ \\
                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $
                \end{tabular} 
            \end{center}
            This method is significantly more computationally intensive. 
    \end{itemize}

\subsection{Binary Knockoffs}
    Though the ``original" knockoff method, described above, achieves the exact desired knockoff properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$. In particular, these values will often have very different empirical distributions. This effect is particularly noticeable when the random variable $x_j$ is discrete but $\tilde X_j$ does not even consist of the same set of values. The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics as linear regression which knockoffs are designed to hit, the original knockoffs will fail to control FDR. \par
    Thus, this paper offers a new method of generating knockoffs which should offer superior performance with other GLMs. In this new method, the matrix of knockoff features $\tilde X$ will be generated by row from a random vector valued variable $\mathbf{\tilde x}$, each entry $\tilde x_j$ of which will have the property $\tilde x_j \sim x_j$. This random variable will also abide by a relaxation of the original knockoff condition:
    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
    This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
    Generating such knockoffs for a completely general case of any $\mathbf x$ is likely infeasible. However, in the specific case where $\mathbf x$ is a random binary vector variable, this paper will demonstrate a method to generate such knockoff variables. Since binary data is extremely common in many areas, this will hopefully serve as a useful tool.

@@

\subsection{Generating Binary Knockoffs}
This theory provides the tools to generate random binary knockoffs. This is done is done sequentially as two steps. The first of these is to come up with the target correlation matrix $C$ for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.

the first of which is to come up with the target covariance matrix for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$. Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and 

@@

        Additionally though, the moment condition must hold for this to correspond to a proper random binary vector. If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross moment matrix of the joint distribution is 

@@

        However, since $\Sigma$ and $\mathbf m$ are from an actual random binary vector, the entries of $\Sigma$ satisfy this property by default. Only off diagonal entries which have had $\mathbf s$ subtracted from them could possibly violate this condition. Furthermore, for these entries, the goal is to make $\Sigma_{L,ij}$ as close to $0$ as possible. Since the condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$, this implies that this method should normally be satisfied. For the SDP method though, the restriction can be built directly into the optimization problem, yielding the updated problem

>>>>

\subsection{Knockoff Filter}
    Let us consider the consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
    \[\mathbf{y} = X\beta + \mathbf{z}\]
    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise. It is important to note that a intercept vector $\mathbf 1$ is not included in $X$. For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$. I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$. The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$. In the situation where $\beta$ is sparse and there is reason to expect that $x_j$ and $y$ are uncorrelated for many $j$, the knockoff filter is a method to select a subset of the $x_j$ which are likely correlated with $y$. \par
    The method specifically seeks to control the false discovery rate (FDR), which is the expected portion of some selection of $x_j$ which are, in fact, uncorrelated with $y$. For a procedure which chooses a subset of variables $\hat S\subseteq \{1,\ldots,p\}$, this is defined as  
    \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: \beta_j=0 \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
    FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$. \par

    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being less correlated with $y$. In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$, the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized. In other words, where $ G := X^TX$, 
    \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}\]
    for some vector $s\in \R^p$ such that, where $\diag{G}$ is the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small. Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$. In the case where $X_j$ is uncorrelated with $\mathbf y$, $\tilde X_j$ will also be uncorrelated with $\mathbf y$. How these knockoffs are actually generated will be described shortly.\par

@@

    In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second. Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor. On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature. This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$. \par

@@

\subsection{Original Knockoff Features}
    The original formulation of the method offers two similar methods of constructing knockoffs. Both of these will, by construction, have exactly the property that
    \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}.\]
    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $X_L$ is
    \[ G_L= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
    $A$, the Schur complement of $G$ in $G_L$ can be calculated as 
    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}G^{-1}\diag{\mathbf s} \]
    For $G_L$ to exist, $G_L$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper} 
    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2G - \diag{\mathbf s} \succeq 0 \]
    Given this is true, $A$ can be factored as $A=C^TC$. Combining this with a satisfactory $\mathbf s$ and an orthonormal matrix $\tilde U\in\R^{n\times p}$ such that $\tilde U^T X = 0$, a $\tilde X$ fulfilling the desired properties can be calculated as 
    \[ \tilde X = X(I-G^{-1}\,\diag{\mathbf s}) + \tilde UC\]
    As mentioned above, $\mathbf s$ should be chosen so as to make $\diag{G}-\mathbf s$ small. Since each $X_j$ has been normalized, this means that $\diag{G}=\mathbf 1$, so $\mathbf 1 - \mathbf s$ should be minimized in accordance with the restrictions on $\mathbf s$. The two methods differ in how $\mathbf s$ is chosen:
    \begin{itemize}
        \item \textit{Equi-correlated knockoffs}: Each original features is set to have the same correlation with its knockoff by setting $\mathbf s = 2\min\{\lambda_{min}(G),1\}\mathbf 1$. For all simulations in this paper, this was the method used.
        \item \textit{SDP knockoffs}: The $\mathbf s$ which minimizes the average correlation between knockoff and original features can be found via a semi-definite programing problem:
            \begin{center}
                \begin{tabular}{r l}
                    minimize & $\|\mathbf1-\mathbf s\|_1$ \\
                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2G $
                \end{tabular} 
            \end{center}
            This method is significantly more computationally intensive. 
    \end{itemize}

\subsection{Binary Knockoffs}
    Though the ``original" knockoff method, described above, achieves the exact desired knockoff properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$. In particular, these values will often have very different empirical distributions. This effect is particularly noticeable when the random variable $x_j$ is discrete but $\tilde X_j$ does not even consist of the same set of values. The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics as linear regression which knockoffs are designed to hit, the original knockoffs will fail to control FDR. \par
    Thus, this paper offers a new method of generating knockoffs which should offer superior performance with other GLMs. In this new method, the matrix of knockoff features $\tilde X$ will be generated by row from a random vector valued variable $\mathbf{\tilde x}$, each entry $\tilde x_j$ of which will have the property $\tilde x_j \sim x_j$. This random variable will also abide by a relaxation of the original knockoff condition:
    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=G \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = G - \diag{\mathbf s}\]
    This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
    Generating such knockoffs for a completely general case of any $\mathbf x$ is likely infeasible. However, in the specific case where $\mathbf x$ is a random binary vector variable, this paper will demonstrate a method to generate such knockoff variables. Since binary data is extremely common in many areas, this will hopefully serve as a useful tool.

@@

\subsection{Generating Binary Knockoffs}
This theory provides the tools to generate random binary knockoffs. This is done is done sequentially as two steps, the first of which is to come up with the target covariance matrix $\Sigma_L$ for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$. This is extremely similar, though not identical, to the method with the original knockoffs, where $X_L^TX_L$ after normalization was targeted. However, much the same method as before can be used. Let the vector of empirical means of the columns of $X$ is $\mathbf m = \frac{1}{n}X^T\mathbf 1$ and the empirical covariance matrix of $X$ is $\Sigma = \frac{1}{n}X^TX - \mathbf{m}\mathbf{m}^T$. Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and 

@@

    While with the original knockoffs, the result was small values for $X_i^T\tilde X_i$, now the result is low covariance between $x_i$ and $\tilde x_i$.\par
        Additionally though, the moment condition must hold for $\Sigma$ to correspond to a proper random binary vector. If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross moment matrix of the joint distribution is 

@@

        However, since $\Sigma$ and $\mathbf m$ are from an actual random binary vector, the entries of $\Sigma$ satisfy this property by default. Only off diagonal entries which have had $\mathbf s$ subtracted from them could possibly violate this condition. Furthermore, for these entries, the goal is to make $\Sigma_{L,ij}$ as close to $0$ as possible. Since the condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$, this implies that this method should normally be satisfied. This is the principle advantage of targeting low covariance between $X_i$ and $\tilde X_i$ rathervh. For the SDP method though, the restriction can be built directly into the optimization problem, yielding the updated problem
